---
title: "Exploration of the relationship between the uniform and exponential distributions"
author: "Peter Furness"
date: "February 2, 2018"
output: ioslides_presentation
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction 

- A shiny app was created for the analysis
- The basic idea is to show that if you sample a number of points from a uniform distribution and order them by size, then the inter-point distances follow an exponential distribution (at least approximately)
- In the app the user can set the sample size and the number of bins used in histograms
- a Kolomogorov-Smirnov (KS) two-sample test gives a quantitative measure of how close the two distribtions are
- In these slides selected parts of the app HTML output are produced with R Markdown. A link to the full app output is provided
- The slides can be used for teaching students about the exponential properties of the uniform distribution!

## Code for the Shiny app

- The ui and server files can be found in the following GitHub repository https://github.com/peterfurness/DevelopingDataProducts
- A simple monte carlo approach is used to generate the samples
- the ks.test function is used for the two-sample KS test

## Selected output from the Shiny app
#### See https://peterfurness.shinyapps.io/DDP_Course_project


```{r, echo = FALSE}
  set.seed(17467)
	n_unif = 1000
	bins = 20
	random_point = runif(n_unif, 0, n_unif)

	## compute interpoint distances, after sorting
	order_point = random_point[order(random_point)]
	
	inter_point = random_point
	inter_point = 0
	for (i in 1:(n_unif-1)) (inter_point[i] = order_point[i+1]-order_point[i])

	lambda = mean(inter_point)
	exponential = rexp(n_unif, rate=lambda)

	mmax = max(inter_point, exponential)

	bands = (0:bins)/bins*mmax
  par(mfrow=c(2,2))
	hist(inter_point, breaks=bands, col = rgb(0,0,1,0.5))
	hist(exponential, breaks = bands, col = rgb(1,0,0,0.5))

	## now the overlay
	hist(inter_point, breaks=bands, main="Overlay of inter_point and exponential", col = rgb(0,0,1,0.5))
	hist(exponential, breaks = bands, col = rgb(1,0,0,0.5), add=TRUE)
```
	
	
```{r, echo = FALSE}
    kst = ks.test(inter_point, exponential)
    
    cat(paste("KS Test: ", "n_unif = ", n_unif, "D = ", 
              signif(kst$statistic, digits=4), 
            "p-value = ", signif(kst$p.value, digits=4))) 
  

```

## Conclusion

- Running the app demostrates that the distribution of inter-point differences from a sample of uniform values is, at least approximately, exponential
- There are some clear limitations in our analysis. For example, in this approach the inter-point differences will always be bounded above, whereas samples from an exponential distribution are not
- A more sophisticated approach would be needed to explore the full asymptotic properties of the distributions
